---
title: "RDCD Public Data Discovery Navigator"
author: "RDCD Core"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: readable
runtime: shiny
---

## 1. Purpose

This R Markdown file is a runnable Shiny prototype for an RDCD data source discovery tool. The app helps investigators move from a broad research idea to a ranked shortlist of public, registered, credentialed, or controlled data sources.

The prototype is designed to use the Excel workbook `RDCD_Data_Source_Preparation_Template.xlsx`. The minimum viable product uses transparent logic: keyword screening, curated categories, filters, and a simple weighted scoring model. A later version can add an AI-assisted layer for keyword expansion, semantic matching, and plain-language explanations.

## 2. Demo concept

Example demo question:

> Diabetes outcomes in rural South Carolina communities, with interest in social vulnerability, food access, and health care access.

Example user inputs:

- Research question: free-text paragraph
- Keywords: diabetes; rural; South Carolina; social vulnerability; food access; Medicaid
- Filters: category, data type, geography, access level, open-only toggle

Expected output:

- Ranked data source table
- Data source cards with fit rationale
- Match score plot
- Access burden and category summaries
- Exportable data-fit summary for RDCD consultation

## 3. Setup

```{r setup, include=FALSE}
required_pkgs <- c(
  "shiny", "bslib", "readxl", "dplyr", "tidyr", "stringr", "purrr",
  "DT", "ggplot2", "glue", "tibble", "tools"
)

missing_pkgs <- required_pkgs[!vapply(required_pkgs, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing_pkgs) > 0) {
  stop(
    "Please install required packages before running this app: ",
    paste(missing_pkgs, collapse = ", "),
    call. = FALSE
  )
}

lapply(required_pkgs, library, character.only = TRUE)
options(shiny.maxRequestSize = 30 * 1024^2)

DEFAULT_WORKBOOK <- "RDCD_Data_Source_Preparation_Template.xlsx"

`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0) y else x
}

clean_names <- function(x) {
  x |>
    stringr::str_replace_all("[^A-Za-z0-9]+", "_") |>
    stringr::str_replace_all("_+$", "") |>
    stringr::str_replace_all("^_+", "") |>
    tolower()
}

as_text <- function(x) {
  x <- ifelse(is.na(x), "", as.character(x))
  stringr::str_squish(x)
}

split_terms <- function(x) {
  if (is.null(x) || length(x) == 0) return(character())
  x <- paste(x, collapse = " ")
  terms <- stringr::str_split(tolower(x), "[,;\\n\\r]+|\\s+and\\s+|\\s+or\\s+")[[1]]
  terms <- stringr::str_squish(terms)
  terms <- terms[nchar(terms) >= 3]
  unique(terms)
}

# --- Search normalization + synonym expansion -------------------------------

normalize_text <- function(x) {
  x <- ifelse(is.na(x), "", as.character(x))
  x <- tolower(x)
  x <- stringr::str_replace_all(x, "[’`]", "'")
  x <- stringr::str_replace_all(x, "[-_/]", " ")
  x <- stringr::str_replace_all(x, "[^a-z0-9\\s]", " ")
  stringr::str_squish(x)
}

light_stem <- function(word) {
  w <- tolower(as.character(word))
  if (nchar(w) <= 3) return(w)

  if (grepl("ies$", w)) {
    return(sub("ies$", "y", w))
  }

  if (grepl("(sses|xes|zes|ches|shes)$", w)) {
    return(sub("es$", "", w))
  }

  if (grepl("ing$", w) && nchar(w) > 5) {
    return(sub("ing$", "", w))
  }

  if (grepl("ed$", w) && nchar(w) > 4) {
    return(sub("ed$", "", w))
  }

  if (grepl("s$", w) && !grepl("(ss|us|is|os|as|es)$", w) && nchar(w) > 4) {
    return(sub("s$", "", w))
  }

  w
}

normalize_search_term <- function(term) {
  term <- normalize_text(term)
  if (!nzchar(term)) return("")

  tokens <- stringr::str_split(term, "\\s+")[[1]]
  tokens <- tokens[nchar(tokens) > 0]
  tokens <- vapply(tokens, light_stem, character(1))
  tokens <- tokens[nchar(tokens) > 0]

  paste(tokens, collapse = " ")
}

normalize_search_terms <- function(x) {
  if (is.null(x) || length(x) == 0) return(character())
  vapply(x, normalize_search_term, character(1), USE.NAMES = FALSE)
}

parse_search_terms <- function(x) {
  if (is.null(x) || length(x) == 0) return(character())

  raw <- paste(x, collapse = " ")

  quoted <- stringr::str_match_all(raw, '"([^"]+)"')[[1]]
  quoted_terms <- if (nrow(quoted) > 0) quoted[, 2] else character()

  stripped <- stringr::str_replace_all(raw, '"[^"]+"', " ")
  pieces <- stringr::str_split(stripped, "[,;\\n\\r]+|\\s+and\\s+|\\s+or\\s+|\\s+")[[1]]

  terms <- c(quoted_terms, pieces)
  terms <- stringr::str_squish(terms)
  terms <- terms[nchar(terms) >= 2]
  unique(terms)
}

search_lexicon <- list(
  diabetes = c(
    "diabetes", "diabetic", "hba1c", "a1c", "glucose", "glycemia", "glycemic",
    "prediabetes", "type 1 diabetes", "type 2 diabetes", "t1d", "t2d", "insulin"
  ),
  cancer = c(
    "cancer", "oncology", "tumor", "tumour", "neoplasm", "malignancy",
    "metastasis", "screening", "survival", "incidence", "mortality", "recurrence",
    "chemotherapy", "radiation", "seer", "tcga", "tcia", "nci", "gdc"
  ),
  cardiovascular = c(
    "cardiovascular", "heart disease", "stroke", "hypertension", "cardiac",
    "coronary", "myocardial infarction", "heart failure", "atrial fibrillation",
    "afib", "cad"
  ),
  mental_health = c(
    "mental health", "depression", "anxiety", "suicide", "behavioral health",
    "ptsd", "bipolar", "schizophrenia", "psychosis", "psychiatric"
  ),
  substance_use = c(
    "opioid", "drug", "substance", "alcohol", "tobacco", "overdose",
    "substance use", "drug use", "alcohol use", "smoking", "vaping"
  ),
  maternal_child_health = c(
    "maternal", "pregnancy", "pregnant", "birth", "prenatal", "postpartum",
    "infant", "child health", "women's health"
  ),
  food_access = c(
    "food access", "food insecurity", "food environment", "food desert",
    "nutrition access", "healthy food access", "food availability"
  ),
  social_vulnerability = c(
    "social vulnerability", "svi", "vulnerability index", "community vulnerability",
    "neighborhood disadvantage", "deprivation"
  ),
  health_care_access = c(
    "health care access", "healthcare access", "access to care", "provider access",
    "medical access", "care access"
  ),
  rurality = c(
    "rural", "rurality", "nonurban", "non urban", "frontier", "remote"
  ),
  claims_payment = c(
    "claims", "medicare", "medicaid", "payment", "spending", "cost", "billing",
    "expenditure", "insurance"
  ),
  ehr_clinical = c(
    "ehr", "electronic health record", "clinical", "encounter", "laboratory",
    "lab results", "medication", "hospital", "icu"
  ),
  environment_exposure = c(
    "environment", "air pollution", "air quality", "pm2.5", "pm 2.5", "ozone",
    "water", "heat", "exposure", "climate"
  ),
  genomics_molecular = c(
    "genomics", "genetic", "gene expression", "biospecimen", "molecular",
    "transcriptomics", "genotype", "phenotype"
  ),
  imaging = c(
    "imaging", "mri", "ct", "radiology", "dicom", "pathology", "scan"
  ),
  geospatial = c(
    "county", "census tract", "tract", "zip", "zcta", "geospatial", "area level",
    "area-level", "neighborhood", "location"
  )
)

build_synonym_index <- function(dict) {
  index <- list()

  for (group_name in names(dict)) {
    group_terms <- unique(c(group_name, dict[[group_name]]))
    group_terms <- unique(vapply(group_terms, normalize_search_term, character(1)))
    group_terms <- group_terms[nzchar(group_terms)]

    for (term in group_terms) {
      index[[term]] <- unique(c(index[[term]], group_terms))
    }
  }

  index
}

synonym_index <- build_synonym_index(search_lexicon)

# ---------------------------------------------------------------------------

extract_words <- function(x) {
  if (is.null(x) || length(x) == 0) return(character())
  stop_words <- c(
    "the", "and", "for", "with", "from", "this", "that", "into", "data", "source",
    "sources", "study", "research", "health", "using", "want", "need", "look", "about",
    "interest", "interested", "outcome", "outcomes"
  )
  words <- stringr::str_split(tolower(paste(x, collapse = " ")), "[^a-z0-9]+")[[1]]
  words <- words[nchar(words) >= 3]
  unique(words[!words %in% stop_words])
}

concept_dictionary <- list(
  cancer = c(
 "cancer","oncology","tumor","tumour","neoplasm",
 "malignancy","metastasis","metastatic",
 "screening","survival","incidence","mortality",
 "recurrence","chemotherapy","radiation",
 "seer","tcga","tcia","nci","gdc","cancer registry"
),
  diabetes = c("diabetes", "hba1c", "glucose", "glycemia", "prediabetes",
               "type 1", "type 2", "T1D", "T2D", "insulin", "A1C", "HbA1c",
               "glycemic", "hyperglycemia", "hypoglycemia"),
  cardiovascular = c(
    "cardiovascular", "heart disease", "stroke", "hypertension",
    "cardiac", "coronary", "myocardial infarction", "MI", "CHF",
    "heart failure", "coronary artery disease", "CAD", "atrial fibrillation",
    "AFib"
  ),
  mental_health = c(
    "mental health", "depression", "anxiety", "suicide", "behavioral health",
    "PTSD", "bipolar", "schizophrenia", "psychosis", "behavioral health", 
    "psychiatric"
  ),
  sdoh = c(
    "social vulnerability", "poverty", "income", "housing", "food access",
    "svi", "acs"
  )
)

expand_terms <- function(terms) {
  if (is.null(terms) || length(terms) == 0) return(character())

  # Normalize input terms
  normalized <- unique(vapply(terms, normalize_search_term, character(1)))
  normalized <- normalized[nzchar(normalized)]

  expanded <- normalized

  # 1) Expand using the synonym index built from search_lexicon
  for (term in normalized) {
    if (!is.null(synonym_index[[term]])) {
      expanded <- c(expanded, synonym_index[[term]])
    }
  }

  # 2) Expand using the concept_dictionary
  concept_terms <- unlist(
    concept_dictionary[intersect(names(concept_dictionary), normalized)],
    use.names = FALSE
  )

  if (length(concept_terms) > 0) {
    expanded <- c(expanded, normalize_search_terms(concept_terms))
  }

  unique(expanded[nzchar(expanded)])
}


contains_any <- function(text, terms) {
  if (length(terms) == 0) return(rep(FALSE, length(text)))
  purrr::map_lgl(text, function(txt) {
    any(stringr::str_detect(txt, stringr::fixed(terms, ignore_case = TRUE)))
  })
}

count_term_hits <- function(text, terms) {
  if (length(terms) == 0) return(rep(0, length(text)))
  purrr::map_int(text, function(txt) {
    sum(stringr::str_detect(txt, stringr::fixed(terms, ignore_case = TRUE)))
  })
}

find_matched_terms <- function(text, terms, max_terms = 12) {
  if (length(terms) == 0 || is.null(text) || length(text) == 0) return("")

  txt <- tolower(paste(text, collapse = " "))

  hits <- terms[vapply(
    terms,
    function(term) stringr::str_detect(txt, stringr::fixed(term, ignore_case = TRUE)),
    logical(1)
  )]

  paste(head(unique(hits), max_terms), collapse = "; ")
}

safe_col <- function(df, nm, default = "") {
  if (nm %in% names(df)) as_text(df[[nm]]) else rep(default, nrow(df))
}

standardize_access_burden <- function(access_level, access_burden) {
  existing <- as_text(access_burden)
  access <- tolower(as_text(access_level))
  inferred <- dplyr::case_when(
    stringr::str_detect(access, "open/public") & !stringr::str_detect(access, "restricted|controlled|credentialed|purchased") ~ "Low",
    stringr::str_detect(access, "registered public-use|public-use plus restricted|registered/controlled|open plus credentialed|open plus controlled|terms") ~ "Medium",
    stringr::str_detect(access, "controlled|restricted|credentialed|purchased") ~ "High",
    TRUE ~ "Unknown"
  )
  ifelse(nchar(existing) > 0, existing, inferred)
}

infer_data_type <- function(text) {
  dplyr::case_when(
    stringr::str_detect(text, "survey|surveys|questionnaire|household|behavioral risk") ~ "Survey",
    stringr::str_detect(text, "registry|surveillance|incidence|mortality|prevalence|vital statistics") ~ "Registry/surveillance",
    stringr::str_detect(text, "claims|medicare|medicaid|payment|cost|expenditure|insurance|provider|enrollment") ~ "Claims/payment",
    stringr::str_detect(text, "ehr|electronic health record|clinical|hospital|icu|encounter|diagnosis|laboratory|medication") ~ "EHR/clinical",
    stringr::str_detect(text, "cohort|longitudinal|follow-up|retirement|aging") ~ "Cohort/longitudinal",
    stringr::str_detect(text, "county|tract|zip|zcta|geographic|geospatial|census|vulnerability|rural") ~ "Geospatial/area-level",
    stringr::str_detect(text, "environment|air quality|water|hazardous|toxic|exposure|heat") ~ "Environmental",
    stringr::str_detect(text, "genom|genotype|phenotype|gene expression|molecular|transcriptomics|biospecimen") ~ "Genomics/molecular",
    stringr::str_detect(text, "imaging|radiology|pathology|mri|ct|dicom") ~ "Imaging",
    stringr::str_detect(text, "physiologic|ecg|waveform") ~ "Physiologic signals",
    stringr::str_detect(text, "clinical trial|trial registry|intervention|sponsor") ~ "Trial registry",
    stringr::str_detect(text, "synthetic") ~ "Synthetic data",
    TRUE ~ "Other"
  )
}

#7/9 improvement
concept_rules <- tibble::tribble(
  ~concept, ~query_pattern, ~source_pattern, ~boost, ~penalty,

  "Cancer",
  "cancer|oncology|tumou?r|neoplasm|malignan",
  "cancer|oncology|seer|tcga|tcia",
  30, 15,

  "Cardiovascular",
  "heart disease|cardiovascular|stroke|hypertension",
  "heart|cardiovascular|stroke|hypertension|cdc wonder|brfss|nhanes",
  25, 12,

  "Maternal",
  "maternal|pregnan|birth|prenatal|infant",
  "maternal|pregnan|birth|prams|nvss",
  25, 15,

  "Mental Health",
  "mental health|depression|anxiety|suicide",
  "mental health|depression|nsduh|mh-cld",
  25, 12,

  "Substance Use",
  "opioid|drug|substance|alcohol|tobacco|overdose",
  "opioid|drug|substance|teds|nsduh|samhsa",
  25, 12,

  "Imaging",
  "mri|ct|radiology|imaging|brain",
  "mri|ct|radiology|imaging|adni|tcia",
  25, 15,

  "EHR",
  "ehr|electronic health record|clinical|icu|encounter",
  "ehr|clinical|mimic|n3c|all of us",
  25, 15,

  "Claims",
  "medicare|medicaid|claims|payment|spending",
  "cms|resdac|claims|medicare|medicaid|meps",
  25, 10,

  "Environment",
  "air pollution|air quality|ozone|pm2\\.5|environment",
  "epa|tracking|air quality|environment",
  20, 10,

  "SDOH",
  "social vulnerability|poverty|income|housing|food access",
  "acs|svi|county health rankings|food environment",
  20, 8
)

score_concepts <- function(query_text, source_text, rules = concept_rules) {
  query_text <- tolower(query_text)
  source_text <- tolower(source_text)

  boost_total <- 0
  penalty_total <- 0

  for (i in seq_len(nrow(rules))) {
    query_hit <- stringr::str_detect(query_text, stringr::regex(rules$query_pattern[i], ignore_case = TRUE))
    source_hit <- stringr::str_detect(source_text, stringr::regex(rules$source_pattern[i], ignore_case = TRUE))

    if (isTRUE(query_hit) && isTRUE(source_hit)) {
      boost_total <- boost_total + rules$boost[i]
    }

    if (isTRUE(query_hit) && !isTRUE(source_hit)) {
      penalty_total <- penalty_total + rules$penalty[i]
    }
  }

  list(boost = boost_total, penalty = penalty_total)
}

#added for 7/9 fix to improve scoring

load_catalog <- function(path) {
  if (!file.exists(path)) {
    return(tibble::tribble(
      ~record_id, ~data_source, ~official_link_text, ~official_url, ~main_data_type_use, ~raw_access_level,
      "DS-017", "CDC PLACES", "PLACES Data Portal", "https://www.cdc.gov/places/", "County, city, census tract, and ZIP-level chronic disease and health behavior estimates", "Open/public",
      "DS-018", "CDC/ATSDR Social Vulnerability Index - SVI", "SVI Data Download", "https://www.atsdr.cdc.gov/place-health/php/svi/", "County and census tract social vulnerability measures", "Open/public",
      "DS-019", "U.S. Census American Community Survey - ACS", "ACS API", "https://www.census.gov/programs-surveys/acs/data/data-via-api.html", "Demographic, socioeconomic, housing, insurance, education, employment variables", "Open/public",
      "DS-047", "USDA Food Environment Atlas", "USDA Food Environment Atlas", "https://www.ers.usda.gov/data-products/food-environment-atlas", "County food access, food stores, restaurants, assistance programs, local foods", "Open/public",
      "DS-053", "South Carolina DPH Public Health Data", "SC DPH Public Health Data", "https://dph.sc.gov/professionals/public-health-data", "South Carolina public health dashboards, reports, BRFSS, cancer registry, HIV/STD, overdose, vital statistics resources", "Open/public plus restricted components"
    ))
  }

  raw <- readxl::read_excel(path, sheet = "Data_Prep") |>
    dplyr::rename_with(clean_names) |>
    dplyr::mutate(dplyr::across(dplyr::everything(), as_text))

 needed <- c(
  "record_id", "data_source", "official_link_text", "official_url", "main_data_type_use", "raw_access_level",
  "access_level_standardized", "access_burden", "primary_category", "secondary_categories_tags",
  "keywords_for_search", "data_type_standardized", "topic_area_s", "population", "geography_coverage",
  "geographic_unit", "unit_of_analysis", "individual_level_data", "area_level_data", "longitudinal",
  "years_available", "refresh_frequency", "linkage_potential", "south_carolina_relevance",
  "common_variables_measures", "suggested_use_cases", "companion_datasets", "limitations_cautions",
  "data_governance_notes", "rdcd_consultation_notes", "demo_include", "priority", "status",
  "app_search_text"
)
  for (nm in needed) if (!nm %in% names(raw)) raw[[nm]] <- ""
  raw[, needed]
}

prepare_catalog <- function(df) {
  df <- df |>
    dplyr::mutate(
      access_level_standardized = dplyr::if_else(
        nchar(access_level_standardized) > 0,
        access_level_standardized,
        raw_access_level
      ),
      search_text = tolower(paste(
        data_source,
        official_link_text,
        main_data_type_use,
        raw_access_level,
        access_level_standardized,
        primary_category,
        secondary_categories_tags,
        keywords_for_search,
        data_type_standardized,
        topic_area_s,
        population,
        geography_coverage,
        geographic_unit,
        unit_of_analysis,
        common_variables_measures,
        suggested_use_cases,
        companion_datasets,
        limitations_cautions,
        data_governance_notes,
        rdcd_consultation_notes,
        app_search_text,
        sep = " "
      )),
      concept_text = tolower(paste(
        primary_category,
        secondary_categories_tags,
        keywords_for_search,
        common_variables_measures,
        suggested_use_cases,
        companion_datasets,
        limitations_cautions,
        rdcd_consultation_notes,
        sep = " "
      )),

      # Normalized text for forgiving search
    search_text_normalized = normalize_text(search_text),
    concept_text_normalized = normalize_text(concept_text),

    data_source_norm = normalize_search_terms(data_source),
    keywords_for_search_norm = normalize_search_terms(keywords_for_search),
    primary_category_norm = normalize_search_terms(primary_category),
    secondary_categories_tags_norm = normalize_search_terms(secondary_categories_tags),
    suggested_use_cases_norm = normalize_search_terms(suggested_use_cases),
    main_data_type_use_norm = normalize_search_terms(main_data_type_use),
    limitations_cautions_norm = normalize_search_terms(limitations_cautions),
    rdcd_consultation_notes_norm = normalize_search_terms(rdcd_consultation_notes),

      category_for_filter = dplyr::if_else(
        nchar(primary_category) > 0,
        primary_category,
        ""
      ),
      data_type_for_filter = dplyr::if_else(
        nchar(data_type_standardized) > 0,
        data_type_standardized,
        infer_data_type(search_text)
      ),
      access_burden_for_filter = standardize_access_burden(access_level_standardized, access_burden),
      official_url = dplyr::if_else(nchar(official_url) > 0, official_url, NA_character_)
    )

  df
}


score_catalog <- function(df, terms, category_filter, data_type_filter, geography_filter, access_filter, open_only) {
  if (nrow(df) == 0) return(df)

  source_terms <- unique(terms[nzchar(terms)])
  query_text <- normalize_text(paste(terms, collapse = " "))

  field_hits <- function(text, terms) {
    if (length(terms) == 0) return(rep(0, length(text)))
    purrr::map_int(text, function(txt) {
      sum(vapply(
        terms,
        function(term) stringr::str_detect(txt, stringr::fixed(term, ignore_case = TRUE)),
        logical(1)
      ))
    })
  }

  weighted_field_score <- function(text, terms, weight) {
    hits <- field_hits(text, terms)
    pmin(100, hits * weight)
  }

  keyword_hits_vec <- field_hits(df$search_text_normalized, source_terms)

  score_data_source <- weighted_field_score(df$data_source_norm, source_terms, 8)
  score_keywords <- weighted_field_score(df$keywords_for_search_norm, source_terms, 7)
  score_category <- weighted_field_score(df$primary_category_norm, source_terms, 6)
  score_secondary <- weighted_field_score(df$secondary_categories_tags_norm, source_terms, 5)
  score_use_cases <- weighted_field_score(df$suggested_use_cases_norm, source_terms, 3)
  score_main_use <- weighted_field_score(df$main_data_type_use_norm, source_terms, 3)
  score_notes <- weighted_field_score(
    paste(df$limitations_cautions_norm, df$rdcd_consultation_notes_norm),
    source_terms,
    1
  )

  keyword_score <- pmin(
    50,
    score_data_source + score_keywords + score_category + score_secondary +
      score_use_cases + score_main_use + score_notes
  )

  concept_score_vec <- purrr::map_dbl(df$concept_text_normalized, function(src) {
    fit <- score_concepts(query_text, src)
    fit$boost - fit$penalty
  })
  concept_score_vec <- pmax(0, concept_score_vec)
  concept_hit <- concept_score_vec > 0

  category_score <- if (length(category_filter) == 0) {
    rep(0, nrow(df))
  } else {
    ifelse(contains_any(df$category_for_filter, category_filter), 20, 0)
  }

  data_type_score <- if (length(data_type_filter) == 0) {
    rep(0, nrow(df))
  } else {
    ifelse(contains_any(df$data_type_for_filter, data_type_filter), 15, 0)
  }

  geography_text <- tolower(paste(
    df$geography_coverage,
    df$geographic_unit,
    df$main_data_type_use,
    df$data_source,
    sep = " "
  ))
  geography_terms <- geography_filter_terms(geography_filter)
  geography_score <- if (length(geography_filter) == 0) {
    rep(0, nrow(df))
  } else {
    ifelse(contains_any(geography_text, geography_terms), 15, 0)
  }

  access_score <- ifelse(df$access_burden_for_filter == "Low", 3,
                         ifelse(df$access_burden_for_filter == "Medium", 2,
                                ifelse(df$access_burden_for_filter == "High", 1, 0)))

  linkage_score <- ifelse(tolower(df$linkage_potential) == "high", 3,
                          ifelse(tolower(df$linkage_potential) == "medium", 2, 0))

  matched_terms_text <- purrr::map_chr(df$search_text_normalized, ~ find_matched_terms(.x, source_terms))

  raw_score <- keyword_score + concept_score_vec + category_score + data_type_score +
    geography_score + access_score + linkage_score

  out <- df |>
    dplyr::mutate(
      keyword_hits = keyword_hits_vec,
      concept_score = round(concept_score_vec, 1),
      concept_hit = concept_hit,
      matched_terms = matched_terms_text,
      match_score = round(pmax(0, pmin(100, raw_score)), 1),
      why_matched = paste0(
        "Matched terms: ", matched_terms_text,
        ". Concept fit: ", round(concept_score_vec, 1),
        ". Category: ", category_for_filter,
        ". Access: ", access_level_standardized,
        "."
      )
    )

  if (length(access_filter) > 0) {
    out <- out |> dplyr::filter(contains_any(access_level_standardized, access_filter))
  }

  if (isTRUE(open_only)) {
    out <- out |>
      dplyr::filter(
        access_burden_for_filter == "Low" |
          stringr::str_detect(tolower(access_level_standardized), "open|public")
      )
  }

  out |>
    dplyr::arrange(dplyr::desc(match_score), dplyr::desc(keyword_hits), data_source)
}


make_html_link <- function(label, url) {
  ifelse(
    is.na(url) | nchar(url) == 0,
    label,
    sprintf('<a href="%s" target="_blank">%s</a>', url, label)
  )
}

make_limitations_tooltip <- function(x) {
  x <- as_text(x)

  purrr::map_chr(x, function(txt) {
    if (!nzchar(txt)) {
      return('<span class="limitations-empty">Not specified</span>')
    }

    # Prevent quotes or HTML in the workbook from breaking the attribute.
    safe_txt <- htmltools::htmlEscape(txt, attribute = TRUE)

    sprintf(
      paste0(
        '<span ',
        'class="limitations-trigger" ',
        'tabindex="0" ',
        'data-bs-toggle="tooltip" ',
        'data-bs-placement="bottom" ',
        'title="%s">',
        'View limitations',
        '</span>'
      ),
      safe_txt
    )
  })
}

DEFAULT_CATEGORIES <- c(
  "Chronic disease",
  "Cancer",
  "Aging",
  "Maternal/child health",
  "Mental health",
  "Substance use",
  "HIV/infectious disease",
  "Health care utilization/cost",
  "Claims/payment/provider",
  "EHR/clinical",
  "Genomics/molecular",
  "Imaging",
  "Clinical trials",
  "SDOH/geospatial",
  "Environmental exposure",
  "Food/nutrition",
  "Rurality",
  "South Carolina/local data",
  "Global health",
  "Synthetic/methods",
  "Other"
)

DEFAULT_DATA_TYPES <- c(
  "Survey", "Registry/surveillance", "Claims/payment", "EHR/clinical", "Cohort/longitudinal",
  "Geospatial/area-level", "Environmental", "Genomics/molecular", "Imaging", "Physiologic signals",
  "Trial registry", "Synthetic data", "Reports/dashboard", "Other"
)
DEFAULT_GEO <- c("United States", "South Carolina", "Global/international", "State-level", "County-level", "Census tract", "ZIP/ZCTA", "City", "Hospital/provider")
DEFAULT_ACCESS <- c("Open/public", "Registered public-use", "Registered/controlled", "Credentialed access", "Controlled/restricted", "Public-use plus restricted files", "Open plus controlled data", "Open plus credentialed access", "Open/public with terms")

normalize_geography_token <- function(x) {
  x <- tolower(as_text(x))
  x <- stringr::str_squish(x)
  if (length(x) == 0 || nchar(x) == 0 || x %in% c("other", "unknown", "n/a", "na")) return(character())

  dplyr::case_when(
    stringr::str_detect(x, "south carolina|^sc$") ~ "South Carolina",
    stringr::str_detect(x, "global|international|country") ~ "Global/international",
    stringr::str_detect(x, "united states|^us$|^u\\.s\\.?$|national") ~ "United States",
    stringr::str_detect(x, "census tract|tract") ~ "Census tract",
    stringr::str_detect(x, "block group") ~ "Census block group",
    stringr::str_detect(x, "zip|zcta") ~ "ZIP/ZCTA",
    stringr::str_detect(x, "county") ~ "County-level",
    stringr::str_detect(x, "state") ~ "State-level",
    stringr::str_detect(x, "city") ~ "City",
    stringr::str_detect(x, "hospital|health system|facility|provider") ~ "Hospital/provider",
    TRUE ~ NA_character_
  ) |>
    stats::na.omit() |>
    as.character()
}

geography_filter_choices <- function(df) {
  raw_values <- c(safe_col(df, "geography_coverage"), safe_col(df, "geographic_unit"))
  raw_tokens <- unlist(stringr::str_split(raw_values, "[,;/\\n\\r]+|\\s+and\\s+|\\s+or\\s+"))
  normalized <- unique(unlist(purrr::map(raw_tokens, normalize_geography_token)))
  normalized <- sort(normalized[nchar(normalized) > 0])
  unique(c(DEFAULT_GEO, normalized))
}

geography_filter_terms <- function(filters) {
  if (length(filters) == 0) return(character())
  terms <- purrr::map(filters, function(filter) {
    dplyr::case_when(
      filter == "United States" ~ "united states; us; u.s.; national",
      filter == "South Carolina" ~ "south carolina; sc",
      filter == "Global/international" ~ "global; international; country",
      filter == "State-level" ~ "state; state-level",
      filter == "County-level" ~ "county; county-level",
      filter == "Census tract" ~ "census tract; tract",
      filter == "Census block group" ~ "census block group; block group",
      filter == "ZIP/ZCTA" ~ "zip; zip code; zcta",
      filter == "City" ~ "city",
      filter == "Hospital/provider" ~ "hospital; health system; facility; provider",
      TRUE ~ filter
    )
  }) |>
    unlist() |>
    split_terms()
  unique(terms)
}
```



## 4. App interface

```{r ui, echo=FALSE}

theme <- bs_theme(
  version = 5,
  primary = "#73000A",
  secondary = "#9E1B32",
  bg = "#F7F7F7",
  fg = "#333333"
)

fluidPage(
  theme = theme,

  tags$head(
    tags$style(HTML("
      :root {
        --bs-primary: #73000A;
        --bs-secondary: #9E1B32;
        --bs-link-color: #73000A;
        --bs-link-hover-color: #5B0008;
      }

      .rdcd-app {
        background: #F7F7F7;
        color: #333333;
      }

      .rdcd-shell {
        max-width: 1440px;
        margin: 0 auto;
        padding: 24px 28px 32px;
      }

      .rdcd-title {
        margin: 0 0 6px 0;
        font-size: 32px;
        font-weight: 800;
        line-height: 1.1;
        color: #73000A;
      }

      .rdcd-subtitle {
        margin: 0 0 22px 0;
        font-size: 14px;
        color: #6B7280;
      }

      .rdcd-section {
        margin-bottom: 18px;
      }

      .rdcd-section-label {
        margin: 0 0 8px 0;
        font-size: 12px;
        font-weight: 800;
        letter-spacing: 0.06em;
        text-transform: uppercase;
        color: #73000A;
      }

      .rdcd-details {
        margin-top: 6px;
      }

      .rdcd-details summary {
        cursor: pointer;
        font-weight: 600;
        color: #73000A;
        list-style: none;
      }

      .rdcd-details summary::-webkit-details-marker {
        display: none;
      }

      .rdcd-details-body {
        margin-top: 10px;
      }

      .rdcd-results-header {
        margin: 0 0 12px 0;
      }

      .rdcd-results-header h2 {
        margin: 0;
        font-size: 20px;
        font-weight: 800;
        color: #1F2937;
      }

      .rdcd-results-header p {
        margin: 4px 0 0 0;
        color: #6B7280;
      }

      .rdcd-app .shiny-input-container {
        width: 100% !important;
        max-width: 100% !important;
        min-width: 0 !important;
      }

      .shiny-input-container,
      .form-group {
        margin-bottom: 12px;
      }

      .form-control,
      .selectize-input {
        border-radius: 10px !important;
        border-color: #D1D5DB !important;
      }

      .btn-primary {
        background-color: #73000A !important;
        border-color: #73000A !important;
        border-radius: 10px;
        font-weight: 600;
      }

      .btn-primary:hover,
      .btn-primary:focus,
      .btn-primary:active {
        background-color: #5B0008 !important;
        border-color: #5B0008 !important;
      }

      .tab-content {
        background: transparent;
        border: none;
        box-shadow: none;
        padding: 12px 0 0 0;
      }

      .nav-tabs {
        border-bottom: 1px solid #E5E7EB;
      }

      .nav-tabs > li > a,
      .nav-tabs .nav-link {
        border: none;
        border-bottom: 2px solid transparent;
        border-radius: 0;
        background: transparent;
        margin-right: 10px;
        color: #73000A !important;
        font-weight: 600;
      }

      .nav-tabs > li.active > a,
      .nav-tabs > li.active > a:hover,
      .nav-tabs > li.active > a:focus,
      .nav-tabs .nav-link.active,
      .nav-tabs .nav-item.show .nav-link {
        border-bottom-color: #73000A !important;
        color: #73000A !important;
        background: transparent !important;
      }

      .data-card {
        background: #FFFFFF;
        border: 1px solid #EEF2F7;
        border-left: 4px solid #73000A;
        border-radius: 12px;
        padding: 16px;
        margin-bottom: 12px;
        box-shadow: none;
      }

      .small-note {
        font-size: 90%;
        color: #6B7280;
      }

      .fit-summary-shell {
        display: flex;
        flex-direction: column;
        gap: 12px;
      }

      .fit-summary-hero {
        background: linear-gradient(135deg, #73000A 0%, #5B0008 100%);
        color: #FFFFFF;
        border-radius: 16px;
        padding: 20px 22px;
        box-shadow: none;
      }

      .fit-summary-kicker {
        font-size: 12px;
        font-weight: 700;
        letter-spacing: 0.08em;
        text-transform: uppercase;
        opacity: 0.9;
        margin-bottom: 6px;
      }

      .fit-summary-hero h3 {
        margin: 0 0 8px 0;
        font-size: 24px;
        font-weight: 800;
      }

      .fit-summary-lead {
        margin: 0;
        font-size: 15px;
        opacity: 0.95;
        max-width: 860px;
      }

      .fit-summary-callout {
        background: #FFFFFF;
        border: 1px solid #EEF2F7;
        border-left: 4px solid #73000A;
        border-radius: 14px;
        padding: 16px 18px;
        box-shadow: none;
      }

      .fit-summary-pre {
        margin: 0;
        white-space: pre-wrap;
        word-wrap: break-word;
        font-family: inherit;
        font-size: 15px;
        line-height: 1.6;
        color: #333333;
        background: transparent;
        border: none;
        padding: 0;
      }

      .fit-summary-actions {
        display: flex;
        justify-content: flex-end;
        margin-top: 4px;
      }

      .rdcd-app .well {
        background: transparent;
        border: none;
        box-shadow: none;
        padding: 0;
        margin-bottom: 0;
      }

      .tooltip {
        z-index: 9999 !important;
      }

      .tooltip-inner {
        max-width: 360px !important;
        padding: 10px 12px !important;
        font-size: 12px !important;
        font-weight: 400 !important;
        line-height: 1.45 !important;
        text-align: left !important;
        white-space: normal !important;
      }

      td.limitations-cell {
        width: 125px !important;
        min-width: 125px !important;
        max-width: 125px !important;
        white-space: nowrap;
        vertical-align: top;
      }

      .limitations-trigger {
        display: inline-block;
        color: #73000A;
        cursor: help;
        font-size: 12px;
        font-weight: 700;
        line-height: 1.2;
        white-space: nowrap;
        border-bottom: 1px dotted #73000A;
      }

      .limitations-trigger:hover {
        color: #5B0008;
      }

      .limitations-trigger:focus {
        outline: 2px solid rgba(115, 0, 10, 0.3);
        outline-offset: 3px;
        border-radius: 2px;
      }

      .limitations-empty {
        color: #6B7280;
        font-size: 12px;
        font-style: italic;
      }
    ")),

    tags$script(HTML("
      function initializeRdcdTooltips() {
        if (!window.bootstrap || !bootstrap.Tooltip) return;

        document.querySelectorAll('.limitations-trigger').forEach(function(element) {
          if (!bootstrap.Tooltip.getInstance(element)) {
            new bootstrap.Tooltip(element, {
              container: 'body',
              trigger: 'hover focus',
              boundary: 'document.body',
              customClass: 'limitations-tooltip',
              placement: 'bottom'
            });
          }
        });
      }

      $(document).on('shiny:connected shiny:recalculated draw.dt shown.bs.tab', function() {
        window.setTimeout(initializeRdcdTooltips, 0);
      });
    "))
  ),

  div(
    class = "rdcd-app",
    div(
      class = "rdcd-shell",

      h1(class = "rdcd-title", "RDCD Public Data Discovery Navigator"),
      div(class = "rdcd-subtitle", "University of South Carolina • Research Design & Data Coordination Core"),

      sidebarLayout(
        sidebarPanel(
          width = 4,

          div(
            class = "rdcd-section",
            div(class = "rdcd-section-label", "Search"),
            fileInput("catalog_file", "Workbook", accept = c(".xlsx")),
            p(
              class = "small-note",
              "Upload a completed RDCD Data Source Preparation workbook or use the default workbook."
            )
          ),

          div(
            class = "rdcd-section",
            div(class = "rdcd-section-label", "Research Question"),
            textAreaInput(
              "question",
              NULL,
              value = "",
              rows = 4,
              width = "100%"
            )
          ),

          div(
            class = "rdcd-section",
            div(class = "rdcd-section-label", "Keywords"),
            textInput("keyword_string", "Keywords", value = ""),
            checkboxInput(
              "use_category_expansion",
              "Expand search using category keyword dictionary",
              TRUE
            )
          ),

          tags$details(
            class = "rdcd-details",
            tags$summary("Filters"),
            div(
              class = "rdcd-details-body",
              selectizeInput(
                "category_filter",
                "Category",
                choices = DEFAULT_CATEGORIES,
                multiple = TRUE,
                options = list(placeholder = "Select categories")
              ),
              selectizeInput(
                "data_type_filter",
                "Data Type",
                choices = DEFAULT_DATA_TYPES,
                multiple = TRUE,
                options = list(placeholder = "Select data types")
              ),
              selectizeInput(
                "geography_filter",
                "Geography",
                choices = DEFAULT_GEO,
                multiple = TRUE,
                options = list(placeholder = "Select geography")
              ),
              selectizeInput(
                "access_filter",
                "Access Level",
                choices = DEFAULT_ACCESS,
                multiple = TRUE,
                options = list(placeholder = "Select access levels")
              ),
              checkboxInput(
                "open_only",
                "Show only open/public or low-burden sources",
                FALSE
              ),
              sliderInput(
                "min_score",
                "Minimum Match Score",
                min = 0,
                max = 100,
                value = 15,
                step = 5
              ),
              numericInput(
                "top_n",
                "Number of Results",
                value = 10,
                min = 5,
                max = 50,
                step = 5
              )
            )
          ),

          div(
            class = "rdcd-section",
            div(class = "rdcd-section-label", "Export"),
            downloadButton("download_results", "Download Results")
          )
        ),

        mainPanel(
          width = 8,

          div(
            class = "rdcd-results-header",
            h2("Results"),
            p("Top matches, cards, comparison, and fit summary.")
          ),

          tabsetPanel(
            tabPanel(
              "Top Matches",
              uiOutput("result_summary"),
              DT::DTOutput("results_table")
            ),

            tabPanel(
              "Data Source Cards",
              uiOutput("source_cards")
            ),

            tabPanel(
              "Compare Sources",
              uiOutput("compare_picker"),
              DT::DTOutput("compare_table")
            ),

            tabPanel(
              "Visual Summary",
              plotOutput("score_plot", height = "420px"),
              plotOutput("category_plot", height = "420px")
            ),

            tabPanel(
              "Data Fit Summary",
              div(
                class = "fit-summary-shell",

                div(
                  class = "fit-summary-hero",
                  div(class = "fit-summary-kicker", "AI-assisted recommendation"),
                  h3("Data Fit Recommendation"),
                  p(
                    class = "fit-summary-lead",
                    "A concise summary of the best-fit source(s), access path, and next RDCD step."
                  )
                ),

                uiOutput("fit_summary"),

                div(
                  class = "fit-summary-actions",
                  downloadButton("download_summary", "Download Summary")
                )
              )
            )
          )
        )
      )
    )
  )
)

```


## 5. Server logic

```{r server, context="server"}
current_path <- reactive({
  if (!is.null(input$catalog_file)) {
    input$catalog_file$datapath
  } else {
    DEFAULT_WORKBOOK
  }
})

raw_catalog <- reactive({
  load_catalog(current_path())
})

processed_catalog <- reactive({
  prepare_catalog(raw_catalog())
})

observe({
  df <- processed_catalog()
  category_choices <- unique(c(DEFAULT_CATEGORIES, split_terms(df$category_for_filter)))
  category_choices <- sort(category_choices[nchar(category_choices) > 0])
  data_type_choices <- unique(c(DEFAULT_DATA_TYPES, df$data_type_for_filter))
  data_type_choices <- sort(data_type_choices[nchar(data_type_choices) > 0])
  geography_choices <- geography_filter_choices(df)
  access_choices <- unique(c(DEFAULT_ACCESS, df$access_level_standardized))
  access_choices <- sort(access_choices[nchar(access_choices) > 0])

  updateSelectizeInput(session, "category_filter", choices = category_choices, server = TRUE)
  updateSelectizeInput(session, "data_type_filter", choices = data_type_choices, server = TRUE)
  updateSelectizeInput(session, "geography_filter", choices = geography_choices, server = TRUE)
  updateSelectizeInput(session, "access_filter", choices = access_choices, server = TRUE)
})



search_terms <- reactive({
  question_terms <- unique(c(
    parse_search_terms(input$question),
    extract_words(input$question)
  ))

  keyword_terms <- parse_search_terms(input$keyword_string)

  base_terms <- unique(c(question_terms, keyword_terms))
  base_terms <- base_terms[nchar(base_terms) >= 2]

  if (isTRUE(input$use_category_expansion)) {
    expand_terms(base_terms)
  } else {
    unique(vapply(base_terms, normalize_search_term, character(1)))
  }
})

scored_results <- reactive({
  score_catalog(
    df = processed_catalog(),
    terms = search_terms(),
    category_filter = input$category_filter %||% character(),
    data_type_filter = input$data_type_filter %||% character(),
    geography_filter = input$geography_filter %||% character(),
    access_filter = input$access_filter %||% character(),
    open_only = input$open_only
  ) |>
    dplyr::filter(match_score >= input$min_score)
})
output$result_summary <- renderUI({
  df <- scored_results()
  tagList(
    tags$p(
      tags$span(class = "metric", nrow(df)), " data sources meet the current criteria. ",
      tags$span(class = "metric", length(search_terms())), " search terms used."
    ),
    tags$p(class = "small-note", paste("Search terms:", paste(search_terms(), collapse = ", ")))
  )
})

output$results_table <- DT::renderDT({
  df <- scored_results() |>
    dplyr::slice_head(n = input$top_n) |>
    dplyr::mutate(
      Rank = dplyr::row_number(),
      link = make_html_link(official_link_text, official_url)
    ) |>
    dplyr::select(
  Rank,
  `Data Source` = data_source,
  `Match Score` = match_score,
  `Access` = access_level_standardized,
  `Category` = category_for_filter,
  `Data Type` = data_type_for_filter,
  `Official Link` = link
)

  DT::datatable(
    df,
    escape = FALSE,
    rownames = FALSE,
    filter = "top",
    options = list(pageLength = 10, scrollX = TRUE, autoWidth = TRUE)
  )
})

output$source_cards <- renderUI({
  df <- scored_results() |> dplyr::slice_head(n = min(10, input$top_n))
  if (nrow(df) == 0) return(tags$p("No data sources meet the current criteria."))

  tagList(purrr::map(seq_len(nrow(df)), function(i) {
    row <- df[i, ]

    tags$div(
      class = "data-card",

      tags$h4(paste0(i, ". ", row$data_source, " - ", row$match_score, "% match")),

      tags$p(
        tags$b("Access: "), row$access_level_standardized,
        " | ",
        tags$b("Burden: "), row$access_burden_for_filter
      ),

      tags$p(tags$b("Category: "), row$category_for_filter),
      tags$p(tags$b("Data type: "), row$data_type_for_filter),
      tags$p(tags$b("Main use: "), row$main_data_type_use),

      tags$details(
        class = "rdcd-details",
        tags$summary("Why it matched"),
        div(
          class = "rdcd-details-body",
          tags$p(row$why_matched)
        )
      ),

      if (!is.na(row$official_url) && nchar(row$official_url) > 0) {
        tags$a(href = row$official_url, target = "_blank", "Open official source")
      } else {
        tags$span(class = "small-note", "No URL provided in the source file.")
      }
    )
  }))
})

output$compare_picker <- renderUI({
  df <- scored_results()
  selectizeInput(
    "compare_sources",
    "Choose 2-5 sources to compare",
    choices = df$data_source,
    selected = head(df$data_source, 3),
    multiple = TRUE
  )
})

output$compare_table <- DT::renderDT({
  selected <- input$compare_sources %||% character()

  validate(
    need(
      length(selected) >= 2,
      "Choose at least two sources to compare."
    ),
    need(
      length(selected) <= 5,
      "Choose no more than five sources."
    )
  )

  df <- scored_results() |>
    dplyr::filter(data_source %in% selected) |>
    dplyr::mutate(
      limitations_tooltip = make_limitations_tooltip(
        limitations_cautions
      )
    ) |>
    dplyr::select(
      `Data Source` = data_source,
      `Match Score` = match_score,
      `Access` = access_level_standardized,
      `Access Burden` = access_burden_for_filter,
      `Category` = category_for_filter,
      `Data Type` = data_type_for_filter,
      `Population` = population,
      `Geography` = geography_coverage,
      `Unit` = unit_of_analysis,
      `Years` = years_available,
      `Linkage Potential` = linkage_potential,
      `SC Relevance` = south_carolina_relevance,
      `Limitations` = limitations_tooltip
    )

  limitations_column <- which(names(df) == "Limitations") - 1L

  DT::datatable(
    df,
    rownames = FALSE,

    # Escape all columns except the tooltip HTML column.
    escape = setdiff(names(df), "Limitations"),

    options = list(
      pageLength = 5,
      scrollX = TRUE,
      autoWidth = FALSE,

      columnDefs = list(
        list(
          targets = limitations_column,
          width = "125px",
          className = "limitations-cell",
          orderable = FALSE
        )
      )
    )
  )
})

output$score_plot <- renderPlot({
  df <- scored_results() |>
    dplyr::slice_head(n = min(15, input$top_n))
  validate(need(nrow(df) > 0, "No results to plot."))
  ggplot(df, aes(x = reorder(data_source, match_score), y = match_score)) +
    geom_col() +
    coord_flip() +
    labs(x = NULL, y = "Match score", title = "Top data source matches") +
    theme_minimal(base_size = 12)
})

output$category_plot <- renderPlot({
  df <- scored_results() |>
    tidyr::separate_rows(category_for_filter, sep = ";") |>
    dplyr::mutate(category_for_filter = stringr::str_squish(category_for_filter)) |>
    dplyr::filter(nchar(category_for_filter) > 0) |>
    dplyr::count(category_for_filter, sort = TRUE) |>
    dplyr::slice_head(n = 15)
  validate(need(nrow(df) > 0, "No categories to plot."))
  ggplot(df, aes(x = reorder(category_for_filter, n), y = n)) +
    geom_col() +
    coord_flip() +
    labs(x = NULL, y = "Number of matching sources", title = "Category distribution among current results") +
    theme_minimal(base_size = 12)
})

data_fit_summary <- reactive({
  df <- scored_results() |> dplyr::slice_head(n = 6)
  if (nrow(df) == 0) return("No data sources meet the current criteria. Try broadening keywords or lowering the minimum score.")

  top <- df[1, ]
  companions <- df |> dplyr::slice(2:min(6, nrow(df))) |> dplyr::pull(data_source)
  companion_text <- if (length(companions) > 0) paste(companions, collapse = "; ") else "None identified in the current search."

  paste0(
    "RDCD DATA FIT SUMMARY\n",
    "=====================\n\n",
    "Research question / concept:\n", input$question, "\n\n",
    "Keywords used:\n", paste(search_terms(), collapse = ", "), "\n\n",
    "Recommended primary source:\n", top$data_source, " (", top$match_score, "% match)\n",
    "Access level: ", top$access_level_standardized, "\n",
    "Access burden: ", top$access_burden_for_filter, "\n",
    "Category: ", top$category_for_filter, "\n",
    "Data type: ", top$data_type_for_filter, "\n",
    "Main use: ", top$main_data_type_use, "\n\n",
    "Suggested companion sources:\n", companion_text, "\n\n",
    "Why this source fits:\n", top$why_matched, "\n\n",
    "Suggested next RDCD step:\n",
    "Confirm the outcome, exposure, population, geography, years, access requirements, and analytic unit before moving to analysis planning or formal data request.\n\n",
    "Caution:\n",
    "This tool supports triage and feasibility review. It does not replace IRB, DUA, data governance, or source-specific access review."
  )
})

output$fit_summary <- renderUI({
  txt <- data_fit_summary()

  if (is.null(txt) || !nzchar(txt)) {
    return(div(class = "fit-summary-callout", p("No summary available yet.")))
  }

  div(
    class = "fit-summary-shell",

    div(
      class = "fit-summary-hero",
      div(class = "fit-summary-kicker", "AI-assisted recommendation"),
      h3("Data Fit Recommendation"),
      p(
        class = "fit-summary-lead",
        "A concise summary of the best-fit source(s), access path, and next RDCD step."
      )
    ),

    div(
      class = "fit-summary-callout",
      tags$pre(class = "fit-summary-pre", txt)
    )
  )
})

output$download_results <- downloadHandler(
  filename = function() {
    paste0("rdcd_data_source_matches_", Sys.Date(), ".csv")
  },
  content = function(file) {
    scored_results() |>
      dplyr::select(
        record_id, data_source, match_score, official_url, main_data_type_use,
        access_level_standardized, access_burden_for_filter, category_for_filter,
        data_type_for_filter, geography_coverage, why_matched
      ) |>
      utils::write.csv(file, row.names = FALSE)
  }
)

output$download_summary <- downloadHandler(
  filename = function() {
    paste0("rdcd_data_fit_summary_", Sys.Date(), ".txt")
  },
  content = function(file) {
    writeLines(data_fit_summary(), file)
  }
)
```


### AI/visualization phase after MVP

After the rule-based prototype works, add:

- Synonym expansion for investigator terms
- Semantic search over data source narratives
- Auto-generated "why this matched" text
- Suggested companion datasets
- Exportable consultation summary
- A chart or matrix showing match score, access burden, geography, and data type
